LDA exploration

### 2016-01-28 doc/lda.Rmd
Added seer to LDA + compared to later visits

```{r ldaData}
lda_ds <- ds %>%
    filter(VN == 0) %>%
    mutate(BCF = rstatsToolkit::aide.tertile(lISSI2) %>%
               factor(., labels = c('1st tertile', '2nd tertile', '3rd tertile'), ordered = TRUE)) %>%
    select(BCF, matches('^ne\\d\\d')) %>%
    mutate_each(funs(as.numeric(scale(.))), -BCF) %>%
    na.omit() %>% 
    arrange(BCF)

# include also the 3-yr and 6-yr data for training/testing.

#fit <- qda(BCF ~ ., data = lda_ds)
fit <- lda(BCF ~ ., data = lda_ds)
fit.cv <- lda(BCF ~ ., data = lda_ds, CV = TRUE)
prop_lda <- round(fit$svd^2/sum(fit$svd^2) * 100, 2)

# proportion explained by each LD
prop_lda
```

```{r testingLDA}
fit.class <- predict(fit, newdata = lda_ds[, -1])$class %>% 
    factor(., ordered = TRUE)
table(fit.class, lda_ds[[1]])
table(fit.cv$class, lda_ds[[1]])
```

```{r FAperBCF}
p_data <- fit$means %>%
    as.data.frame() %>%
    add_rownames('Group') %>%
    gather(FA, Mean,-Group) %>%
    mutate(Group = Group %>% factor(., c('1st tertile', '2nd tertile', '3rd tertile'), ordered = TRUE)) %>%
    arrange(Group, Mean) %>%
    mutate(FA = paste0(Group, FA) %>%
               factor(., unique(.)))
FA.orig <- p_data$FA %>% 
    as.character()
FA.label <- p_data$FA %>%
    gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
    gsub('n(\\d)', 'n-\\1', .) %>%
    as.character()
p_data %>%
    ggplot(aes(x = Mean, y = FA)) +
    geom_point() +
    geom_vline(xintercept = 0) +
    facet_wrap(~ Group, scale = 'free_y') +
    scale_y_discrete(breaks = FA.orig, label = FA.label) +
    ggthemes::theme_tufte(12, 'sans') +
    theme(panel.margin = unit(2, 'lines'),
          panel.grid = element_line(colour = 'black'),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_blank())

```

```{r FAperLD}
p_data2 <- fit$scaling %>%
    as.data.frame() %>%
    add_rownames('Var') %>%
    gather(LD, Value, -Var) %>%
    arrange(LD, Value) %>% 
    mutate(Var = factor(paste0(LD, Var), unique(paste0(LD, Var))))
FA.orig <- p_data2$Var %>% 
    as.character()
FA.label <- p_data2$Var %>%
    gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
    gsub('n(\\d)', 'n-\\1', .) %>%
    as.character()
p_data2 %>% 
    ggplot(aes(x = Value, y = Var)) +
    geom_point() +
    geom_vline(xintercept = 0) +
    facet_wrap(~LD, scale = 'free_y') +
    scale_y_discrete(breaks = FA.orig, label = FA.label) +
    ggthemes::theme_tufte(12, 'sans') +
    theme(panel.margin = unit(2, 'lines'),
          panel.grid = element_line(colour = 'black'),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_blank())
```

```{r LDbyBCF}
p_fit <- predict(object = fit, newdata = lda_ds)
# predict on follow up visits too.
p_ds <- data.frame(BCF = lda_ds[,"BCF"], lda = p_fit$x)
ggplot(p_ds, aes(lda.LD1, lda.LD2, group = BCF)) +
    geom_vline(xintercept = 0, linetype = 'dotted') +
    geom_hline(yintercept = 0, linetype = 'dotted') +
    stat_density2d(aes(alpha = ..level.., colour = BCF), size = 2) +
    labs(x = paste0("LD1 (", prop_lda[1], "%)"),
         y = paste0('LD2 (', prop_lda[2], '%)')) +
    ggthemes::theme_tufte(12, 'sans') +
    scale_alpha(guide = 'none') +
    theme(legend.title = element_blank(),
          legend.position = c(1,0),
          legend.justification = c(1,0))
```

>>>>

```{r ldaData}
ds.prep <- ds %>%
    select(VN, lISSI2, matches('^ne\\d\\d')) %>% 
    group_by(VN) %>% 
    mutate(BCF = rstatsToolkit::aide.tertile(lISSI2) %>%
               factor(., labels = c('1st tertile', '2nd tertile', '3rd tertile'), ordered = TRUE)) %>%
    select(BCF, everything()) %>% 
    select(-lISSI2) %>% 
    mutate_each(funs(as.numeric(scale(.))), -BCF, -VN) %>%
    ungroup()

df_Dysgly <- ds %>% 
    filter(VN == 0) %>% 
    mutate(Dysgly = factor(IGT + IFG + DM, labels = c('no', 'yes'))) %>% 
    select(Dysgly, matches('^ne\\d\\d')) %>%
    mutate_each(funs(as.numeric(scale(.))), -Dysgly) %>% 
    na.omit()

df_y1 <- ds.prep %>% 
    filter(VN == 0) %>%
    select(-VN) %>% 
    na.omit() %>%
    arrange(BCF)
df_y3 <- ds.prep %>% 
    filter(VN == 1) %>% 
    select(-VN) %>% 
    na.omit() %>% 
    arrange(BCF)
df_y6 <- ds.prep %>% 
    filter(VN == 2) %>% 
    select(-VN) %>% 
    na.omit() %>% 
    arrange(BCF)

# include also the 3-yr and 6-yr data for training/testing.
fit_y1 <- lda(BCF ~ ., data = df_y1)
fit_y3 <- lda(BCF ~ ., data = df_y3)
fit_y6 <- lda(BCF ~ ., data = df_y6)
fit_dg <- lda(Dysgly ~ ., data = df_Dysgly)
# compare with CV lda.
fitcv_y1 <- lda(BCF ~ ., data = df_y1, CV = TRUE)
fitcv_y3 <- lda(BCF ~ ., data = df_y3, CV = TRUE)
fitcv_y6 <- lda(BCF ~ ., data = df_y6, CV = TRUE)
```

```{r testingLDA}
fit.class_y1 <- predict(fit_y1, newdata = df_y1[, -1])$class %>% 
    factor(., ordered = TRUE)
fit.class_y3 <- predict(fit_y3, newdata = df_y3[, -1])$class %>% 
    factor(., ordered = TRUE)
fit.class_y6 <- predict(fit_y6, newdata = df_y6[, -1])$class %>% 
    factor(., ordered = TRUE)

table(fit.class_y1, df_y1[[1]])
table(fitcv_y1$class, df_y1[[1]])

table(fit.class_y3, df_y3[[1]])
table(fitcv_y3$class, df_y3[[1]])

table(fit.class_y6, df_y6[[1]])
table(fitcv_y6$class, df_y6[[1]])

# Amount of participants who were misclassified
table(fit.class_y1, df_y1[[1]]) %>% 
    as.data.frame() %>% 
    filter(fit.class_y1 != Var2) %>% 
    summarize(count = sum(Freq))

# Amount of participants who were correctly classified
table(fit.class_y1, df_y1[[1]]) %>% 
    as.data.frame() %>% 
    filter(fit.class_y1 == Var2) %>% 
    fit.class_y1 <- predict(fit_y1, newdata = df_y1[, -1])$class %>% 
    factor(., ordered = TRUE)summarize(count = sum(Freq))

fit.class_dg <- predict(fit_dg, newdata = df_Dysgly[, -1])$class %>% 
    factor(., ordered = TRUE)
table(fit.class_dg, df_Dysgly[[1]])
# Not enough sample size for dysglycemia
```

```{r FAperBCF}
seer::visualize(
    fit_y1, 'means', label.rename.fun = function(x)
        x %>%
        gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
        gsub('n(\\d)', 'n-\\1', .)
    ) %>% 
    seer::vision_simple()
seer::visualize(
    fit_y3, 'means', label.rename.fun = function(x)
        x %>%
        gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
        gsub('n(\\d)', 'n-\\1', .)
    ) %>% 
    seer::vision_simple()
seer::visualize(
    fit_y6, 'means', label.rename.fun = function(x)
        x %>%
        gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
        gsub('n(\\d)', 'n-\\1', .)
    ) %>% 
    seer::vision_simple()
```

```{r FAperLD}
seer::visualize(
    fit_y1, 'scaling', label.rename.fun = function(x)
        x %>%
        gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
        gsub('n(\\d)', 'n-\\1', .)
    ) %>% 
    seer::vision_simple()
seer::visualize(
    fit_y3, 'scaling', label.rename.fun = function(x)
        x %>%
        gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
        gsub('n(\\d)', 'n-\\1', .)
    ) %>% 
    seer::vision_simple()
seer::visualize(
    fit_y6, 'scaling', label.rename.fun = function(x)
        x %>%
        gsub('^.*(\\d\\d)(\\d)', '\\1:\\2', .) %>%
        gsub('n(\\d)', 'n-\\1', .)
    ) %>% 
    seer::vision_simple()
```

```{r LDbyBCF}
seer::visualize(fit_y1, orig.data = df_y1, group.name = 'BCF') %>% 
    seer::vision_simple()
seer::visualize(fit_y3, orig.data = df_y3, group.name = 'BCF') %>% 
    seer::vision_simple()
seer::visualize(fit_y6, orig.data = df_y6, group.name = 'BCF') %>% 
    seer::vision_simple()
```
